The problem and promise of scale in animal social multilayer networks

### 2020-02-09 paper/manuscript.Rmd
spiegel refs

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods (e.g. Carter et al. 2015), whereas social
association represents shared space use by members of the same social group
(Figure \@ref(fig:spacetime), Franks et al. 2010). Further, for social
interactions or associations to occur individuals must share space, and thus
have overlapping home ranges. In the context of the social environment, home
range overlap is a relatively coarse scale of sociality across which individuals
share space over coarser spatial and temporal scales (e.g. Piza-Roca et al.
2018). Although the social environment clearly scales spatially and temporally,
it remains unclear whether coarser scales of sociality, including social
association and home range overlap, vary predictably with social interaction
(Castles et al. 2014; Farine 2015). Thus, we define social scale as the type of
social relationship observed that can be explicitly defined and measured (Farine
et al. 2015).

Animal social network analysis is a well-developed tool used to measure the
relationships of individuals and organization of social systems (Krause et al.
2009; Wey et al. 2008; Croft et al. 2008). Social network analysis provides
insight into the structure of social communities and social network properties,
which can influence population dynamics and evolutionary processes
(Pinter-Wollman et al. 2014; Kurvers et al. 2014). Despite the widespread use
and innovation of traditional social network analysis (Webber & Vander Wal
2019), it typically considers a single scale of sociality in a given network
(monolayer), drastically simplifying the potential complexity of animal social
systems (Finn et al. 2019).

Unlike traditional monolayer social networks, multilayer networks explicitly
consider social systems across contexts, including scale-dependent contexts
(Pilosof et al. 2017). Multilayer networks are made up of multiple layers, each
representing (i) different classes of individuals, e.g. male or female; (ii) types 
of behaviours, e.g. grooming, travelling, or foraging; and (iii) spatial areas,
e.g. local or regional, or temporal windows, e.g. daily or seasonal 
(Kivela et al. 2014; Porter 2018). Multilayer networks are relatively novel to studies of animal
behaviour (Finn et al. 2019; Silk et al. 2018), although they have been used to
describe multidimensional human social systems, complex transportation networks,
and neural networks [@Silk_2018]. The role of scale within multilayer
social networks remains unexplored but provides a unique opportunity to develop
novel understanding of variation in the temporal or spatial scales associated
with social systems.

@@

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO: get Bjorneraas ref-->. We did not collar all female
caribou in the herd; however, we assumed that our sample of collared animals was
random.

@@

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question (Carter, Lee, & Marshall, 2015; Castles et al., 2014; Farine, 2015).
Furthermore, it is a common trope that individuals must be associating to
interact, and many studies therefore assume that proximity is proxy for
interacting (Farine, 2015). For instance, baboons need to be in proximity in
order to groom each other. However, animals have certain interactions that do
not require proximity such as long-distance vocalization and olfactory signals
(Carter et al., 2015). With the advancement of biologging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays (Snijders & Naguib, 2017). These varying social scales can have different
fitness repercussions for individuals. Rhesus macaques (*Macaca mulatta*) had
higher survivorship when they had fewer but stronger, more stable social
associates and grooming partners rather than simply more associates (Ellis,
Snyder-Mackler, Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights
that different qualities of social scales have lasting consequences.

@@

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful simplification of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks (Castles et al., 2014). Furthermore, multilayer networks provide a
framework to test hypotheses linking social phenotypes and fitness across
temporal scales. Because of the non-independent nature of social interactions
and associations, especially over time, it is hard to test hypotheses about
social dynamics with a suitable null control (Farine, 2017; Proskurnikov &
Tempo, 2017). Using the framework of multilayer networks, making layers of
different temporal periods provides a way to make appropriate null models to
test questions about the dynamics of an individual’s social network position
over multiple aggregated time periods.

@@

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that
explicity quantify movement paramters could be itnegrated. For example, edges 
could represent measures of dyadic or collective movement based on dynamic 
interaction metrics [@Long_2014]. Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences [@Webber_2018]. In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
[@Chamaillé_Jammes_2008]. Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
(Spiegel et al. 2016). For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour (Spiegel et al. 2018). Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment [@He_2019].

>>>>

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods (e.g. Carter et al. 2015), whereas social
association represents shared space use by members of the same social group
(Figure \@ref(fig:spacetime), Franks et al. 2010). Further, for social
interactions or associations to occur individuals must share space, and thus
have overlapping home ranges. In the context of the social environment, home
range overlap is a relatively coarse scale of sociality across which individuals
share space over coarser spatial and temporal scales (e.g. Piza-Roca et al.
2018). Although the social environment clearly scales spatially and temporally,
it remains unclear whether coarser scales of sociality, including social
association and home range overlap, vary predictably with social interaction
 [@Castles_2014; @Farine2015a]. Thus, we define social scale as the type of
social relationship observed that can be explicitly defined and measured [@Farine_2015a].

Animal social network analysis is a well-developed tool used to measure the
relationships of individuals and organization of social systems (Krause et al.
2009; Wey et al. 2008; Croft et al. 2008). Social network analysis provides
insight into the structure of social communities and social network properties,
which can influence population dynamics and evolutionary processes
(Pinter-Wollman et al. 2014; Kurvers et al. 2014). Despite the widespread use
and innovation of traditional social network analysis (Webber & Vander Wal
2019), it typically considers a single scale of sociality in a given network
(monolayer), drastically simplifying the potential complexity of animal social
systems [@Finn_2019].

Unlike traditional monolayer social networks, multilayer networks explicitly
consider social systems across contexts, including scale-dependent contexts
(Pilosof et al. 2017). Multilayer networks are made up of multiple layers, each
representing (i) different classes of individuals, e.g. male or female; (ii) types 
of behaviours, e.g. grooming, travelling, or foraging; and (iii) spatial areas,
e.g. local or regional, or temporal windows, e.g. daily or seasonal 
(Kivela et al. 2014; Porter 2018). Multilayer networks are relatively novel to studies of animal
behaviour [Silk_2018; @Finn_2019], although they have been used to
describe multidimensional human social systems, complex transportation networks,
and neural networks [@Silk_2018]. The role of scale within multilayer
social networks remains unexplored but provides a unique opportunity to develop
novel understanding of variation in the temporal or spatial scales associated
with social systems.

@@

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following [@Bjørneraas_2010]. 
We did not collar all female caribou in the herd; however, we assumed that our 
sample of collared animals was random.

@@

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Cater_2015; @Farine_2015a].
Furthermore, it is a common trope that individuals must be associating to
interact, and many studies therefore assume that proximity is proxy for
interacting (Farine, 2015). For instance, baboons need to be in proximity in
order to groom each other. However, animals have certain interactions that do
not require proximity such as long-distance vocalization and olfactory signals
[@Carter_2015] With the advancement of biologging technology, we have
the ability to record social interactions that do not require proximity,
particularly long-distance communication networks with devices like microphone
arrays (Snijders & Naguib, 2017). These varying social scales can have different
fitness repercussions for individuals. Rhesus macaques (*Macaca mulatta*) had
higher survivorship when they had fewer but stronger, more stable social
associates and grooming partners rather than simply more associates (Ellis,
Snyder-Mackler, Ruiz-Lambides, Platt, & Brent, 2019). This finding highlights
that different qualities of social scales have lasting consequences.

@@

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful simplification of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a
framework to test hypotheses linking social phenotypes and fitness across
temporal scales. Because of the non-independent nature of social interactions
and associations, especially over time, it is hard to test hypotheses about
social dynamics with a suitable null control (Farine, 2017; Proskurnikov &
Tempo, 2017). Using the framework of multilayer networks, making layers of
different temporal periods provides a way to make appropriate null models to
test questions about the dynamics of an individual’s social network position
over multiple aggregated time periods.

@@

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that
explicity quantify movement paramters could be itnegrated. For example, edges 
could represent measures of dyadic or collective movement based on dynamic 
interaction metrics [@Long_2014]. Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences [@Webber_2018]. In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
[@Chamaillé_Jammes_2008]. Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
[@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment [@He_2019].
